---
title: "Areas Vulnerable to Covid-Related Income and Job Loss"
output: html_document
description: "Project description here"
tags: ["R", "geospatial"]
weight: 3
draft: false
output: html_document
---

NOTES FOR THIS FILE:
ADD EXPLANATIONS AS NOTED.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F, echo = F}
library(tidyverse)
library(leaflet)
```

```{r}
# Read in data
vulnerable <- read_rds("/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/acs/acs_vuln.Rds") # Sean's ACS 2014/18 vulnerable area indicators data
```

# Areas Vulnerable to Covid-Related Income and Job Loss
Write introduction here.

## Data and Methods
Describe data sources and what we did here.

## Findings
Give a brief summary of findings, then describe each output below (explain summary map and individual indicators with their maps, why we selected them, and main takeaways for each map).

### Relative Vulnerability Map
```{r}
#Summary score leaflet
pal <- colorFactor("Oranges", domain = vulnerable$SummaryScore)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>Workers Unemployed:</strong>",
              round(vulnerable$pct_unempl, 2),
              "<br />",
              "<strong>% Part-Time:</strong>",
              round(vulnerable$pct_part, 2), 
              "<br />",
              "<strong>% High School or Less:</strong>",
              round(vulnerable$pct_hsless, 2), 
              "<br />",
              "<strong>$ Median Worker Income:</strong>",
              formatC(vulnerable$medinc, format = "f", big.mark = ",", digits = 0),
              "<br />",
              "<strong>% In Vulnerable Occupations:</strong>",
              round(vulnerable$pct_occup, 2),
              "<br />",
              "<strong>% In Vulnerable Industries:</strong>",
              round(vulnerable$pct_ind, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(SummaryScore),
              fillOpacity = .8, 
              stroke = FALSE,
              label = labels,
              labelOptions = labelOptions(direction = "bottom",
                                          style = list(
                                            "font-size" = "12px",
                                            "border-color" = "rgba(0,0,0,0.5)",
                                            direction = "auto"))) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~vulnerable$SummaryScore,
            title = "Relative<br>Vulnerability", 
            opacity = 1,
            na.label = "Not Available")
```

# Which Areas Already Experience High Unemployment Rates?
```{r}
#Percent In Labor Force Unemployed
pal <- colorQuantile("Blues", domain = vulnerable$pct_unempl, probs = seq(0, 1, length = 6), right = FALSE)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>Workers Unemployed:</strong>",
              round(vulnerable$pct_unempl, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(pct_unempl),
              fillOpacity = .8, 
              stroke = FALSE, 
              label = labels) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~pct_unempl,
            title = "Percent Unemployed<br>by Quintile Group", 
            opacity = 1,
            na.label = "Not Available",
            labFormat = function(type = "quantile", cuts = 5, p = pct_unempl) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 2), " &ndash; ", round(cuts[-1], 2), ")")
                  })
```

# Where are Part-Time Workers?
```{r}
# Percent In Labor Force Employed Less Than Year Round Full Time
pal <- colorQuantile("Oranges", domain = vulnerable$pct_part, probs = seq(0, 1, length = 6), right = FALSE)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>% Part-Time:</strong>",
              round(vulnerable$pct_part, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(pct_part),
              fillOpacity = .8, 
              stroke = FALSE, 
              label = labels) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~pct_part,
            title = "Percent Employed Part Time<br>by Quintile Group", 
            opacity = 1,
            na.label = "Not Available",
            labFormat = function(type = "quantile", cuts = 5, p = pct_part) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 2), " &ndash; ", round(cuts[-1], 2), ")")
                  })
```

# Where are Workers with High School Education or Less?
```{r}
# Percent in Labor Force with High School Education or Less
pal <- colorQuantile("Blues", domain = vulnerable$pct_hsless, probs = seq(0, 1, length = 6), right = FALSE)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>% High School or Less:</strong>",
              round(vulnerable$pct_hsless, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(pct_hsless),
              fillOpacity = .8, 
              stroke = FALSE, 
              label = labels) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~pct_hsless,
            title = "Percent Workers<br>with HS Education or Less<br>by Quintile Group", 
            opacity = 1,
            na.label = "Not Available",
            labFormat = function(type = "quantile", cuts = 5, p = pct_hsless) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 2), " &ndash; ", round(cuts[-1], 2), ")")
                  })
```

### Where are Low-Income Workers?
```{r}
# Median Worker Income
pal <- colorQuantile("Oranges", domain = vulnerable$medinc, probs = seq(0, 1, length = 6), right = FALSE)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<strong>$ Median Worker Income:</strong>",
              formatC(vulnerable$medinc, format = "f", big.mark = ",", digits = 0)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(medinc),
              fillOpacity = .8, 
              stroke = FALSE, 
              label = labels) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~medinc,
            title = "Median Worker Income<br><br>by Quintile Group", 
            opacity = 1,
            na.label = "Not Available",
            labFormat = function(type = "quantile", cuts = 5, p = medinc) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 2), " &ndash; ", round(cuts[-1], 2), ")")
                  })
```

### Where are Workers in Occupations Vulnerable to Covid-Related Job and Income Loss?
```{r}
# Percent Employed in Occupations Vulnerable to Covid-Related Job and Income Loss
pal <- colorQuantile("Blues", domain = vulnerable$pct_occup, probs = seq(0, 1, length = 6), right = FALSE)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>% In Vulnerable Occupations:</strong>",
              round(vulnerable$pct_occup, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(pct_occup),
              fillOpacity = .8, 
              stroke = FALSE, 
              label = labels) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~pct_occup,
            title = "Percent Employed in<br>Vulnerable Occupations<br>by Quintile Group", 
            opacity = 1,
            na.label = "Not Available",
            labFormat = function(type = "quantile", cuts = 5, p = pct_occup) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 2), " &ndash; ", round(cuts[-1], 2), ")")
                  })
```

### Where are Workers in Industries Vulnerable to Covid-Related Job and Income Loss?
```{r}
# Percent Employed in Industries Vulnerable to Covid-Related Job and Income Loss
pal <- colorQuantile("Oranges", domain = vulnerable$pct_ind, probs = seq(0, 1, length = 6), right = FALSE)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>% In Vulnerable Industries:</strong>",
              round(vulnerable$pct_ind, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(pct_ind),
              fillOpacity = .8, 
              stroke = FALSE, 
              label = labels) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~pct_ind,
            title = "Percent Employed in<br>Vulnerable Industries<br>by Quintile Group", 
            opacity = 1,
            na.label = "Not Available",
            labFormat = function(type = "quantile", cuts = 5, p = pct_ind) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 2), " &ndash; ", round(cuts[-1], 2), ")")
                  })
```

# References
Citations here.