---
title: "Emerging Employment Areas in Fairfax County"
output: html_document
description: "Project description here"
tags: ["R", "geospatial"]
weight: 4
draft: false
output: html_document
---

NOTES FOR THIS FILE:
* COLOR SCHEME
* ADD CONTENT AS NOTED

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(leaflet)
library(dplyr)
library(readxl)
library(readr)
library(sf)
```

```{r setup, include=FALSE}
# Get data
lodespatial <- read_rds("/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/lodes/lodes_spatial.Rds") # Sarah's LODES 2017 workplace + ACS spatial
fortune_geo <- read_sf("/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/fortune/fortune_1000/fairfax_fortune_1000_geo.shp") # Fortune 1000

employers <- read_excel('/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/original/vworks/allemployers.xlsx') %>%
  select_all(~gsub("\\s+|\\.", "_", .)) %>% 
  select_all(tolower) %>% 
  arrange(desc(estimate_of_employees)) %>% 
  slice_head(n = 100) # Susweta's top 100 largest employers
```

# Emerging Employment Areas in Fairfax County
Write introduction here.

## Data and Methods
Describe data sources and what we did here.

## Findings
Give a brief summary of findings, then describe each output below.


### Where and Who are Fairfax County's Largest Employers?
```{r}
pal <- colorQuantile("Reds", employers$estimate_of_employees, probs = seq(0, 1, length = 6), right = FALSE)
labels <- lapply(
  paste("<strong>Company name: </strong>",
         employers$business_name,
        "<br />",
        "<strong>6 digit title: </strong>",
        employers$`6_digit_title`,
        "<br />",
        "<strong>Number of employees: </strong>",
        employers$estimate_of_employees),
        htmltools::HTML
  )
leaflet(data = employers, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(data = employers, lng = employers$longitude, lat = employers$latitude, 
                   radius = 3.5, color = ~pal(estimate_of_employees), 
                   stroke = FALSE, fillOpacity = 1,
                   label = ~labels, labelOptions = labelOptions(noHide = F, direction = "bottom")) %>%
  addLegend("bottomleft", pal = pal, values = employers$estimate_of_employees, 
            title = "Estimated Number of Employees<br>by Quantile Group", opacity = 1, 
            labFormat = function(type = "quantile", cuts = 5, p = data1$estimate_of_employees) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
                  })
```

### Do Fortune 1000 Companies Attract Workers?
Describe how the map works, what it shows, main takeaways.
```{r}
# Fortune 1000 HQ plot
pal <- colorQuantile("Oranges", domain = lodespatial$perTotalJobs, probs = seq(0, 1, length = 6), right = FALSE)

leaflet(data = lodespatial, 
        options = leafletOptions(minZoom = 10))%>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(perTotalJobs), fillOpacity = .7, stroke = FALSE) %>%
  addCircleMarkers(data = fortune_geo, popup = fortune_geo$Name, 
                   radius = 3.5,
                   color = "#232D4B",
                   stroke = FALSE, fillOpacity = 1) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~perTotalJobs,
            title = "Percent of All<br>Fairfax County Jobs,<br>by Quintile Group", 
            opacity = 1,
            labFormat = function(type = "quantile", cuts = 5, p = perTotalJobs) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
                  })
```

### How Many Workers do the Top 100 Employers Attract?
Describe how the map works, what it shows, main takeaways.
```{r}
pal <- colorQuantile("Oranges", domain = lodespatial$perTotalJobs, probs = seq(0, 1, length = 6), right = FALSE)

labels <- lapply(
  paste("<strong>Company name:</strong>",
         employers$business_name,
        "<br />",
        "<strong>Sector:</strong>",
        employers$`6_digit_title`,
        "<br />",
        "<strong>Estimated number of employees:</strong>",
        employers$estimate_of_employees),
        htmltools::HTML
  )

leaflet(data = lodespatial,
        options = leafletOptions(minZoom = 10))%>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(fillColor = ~pal(perTotalJobs), fillOpacity = .7, stroke = FALSE) %>%
  addCircleMarkers(data = employers, label = ~labels, labelOptions = labelOptions(noHide = F, direction = "bottom"),
                   radius = 3.5,
                   color = "#232D4B",
                   stroke = FALSE, fillOpacity = 1) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~perTotalJobs,
            title = "Percent of All<br>Fairfax County Jobs,<br>by Quintile Group", 
            opacity = 1,
            labFormat = function(type = "quantile", cuts = 5, p = perTotalJobs) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
                  })
```

# References
Citations here.