---
title: "Fairfax County Networks"
description: "Project description here"
tags: ["R", "geospatial"]
weight: 5
draft: false
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidycensus)
library(leaflet)
library(sf)
library(shiny)
library(fabricatr)
library(RColorBrewer)
```

# Fairfax County Networks
Write introduction here.

## Data and Methods

The LODES data is produced by the US Census Bureau in order to better understand where workers reside and work. There are three main datasets of the LODES data; we use the Origin-Destination files for our networks. These datasets are available for 2002-2017, but we only use the 2010-2017 datasets.

The OD datasets indicate where individuals work and reside by Census Block Code. We aggregate the data by Census Tract.  





### Interpretation of Centrality Measures:

_Weighted Out-Degree Centrality:_ These plots highlight census tracts from which individuals depart to get to their jobs in other tracts.

_Weighted In-Degree Centrality:_ These plots highlight census tracts that are work destinations.

_Betweenness Centrality:_ Betweenness centrality measures if a particular tract lies on paths between other tracts. 

_Eigenvector Centrality:_ Also known as rank centrality, measures importance of a tract based on if neighboring connections are important, in other words, have lots of connections.


and our explanations of the cent measures should be relatively commonsense. in-degree = in-flows of jobs, out-degree = out-flows of jobs, betweenness = tracts that serve as bridges for workers from disparate tracts to flow through and eigenvector = measure that indicates which tracts that are more important based on they are connected to other well-connected tracts

nice. not sure what to do with eigenvector centrality yet. that essentially seems to reflect weighted in-degree which is not at all surprising

i guess it just means that those high-traveled areas take on more travel flow and infrastructure stress than all other areas. not really that intersting.

## Findings

the findings are a little tougher. i think the in and out-degree are relatively straight-forward. we see workers from the county shift from west to east which tells us about (travel flows at certain times of days) as well as which areas of the county are going to be most busy at certain times of days and how this could impact infrastructure, etc.

the betweenness seems to reflect "flow between" areas which the county could market to emerging industries or other shops as areas that could be easily accessible for workers that are traveling to and from home and work. one could argue that the betweenness centrality of certain tracts could be an indicator of emerging areas for new businesses to setup shop. you might need to take a closer look at this, but when i look at this map from our gentrification work it looks to map closely on to the betweenness centrality map you created it. thus, this would validate that network analysis helps provide some capacity of predicting emerging economic geographies and, in turn, further indicates that infrastructure will need to be built to reflect these emerging trends.


Give a brief summary of findings, then describe each output below (explain summary map and individual indicators with their maps, why we selected them, and main takeaways for each map).

### Initial Unemployment Filings



```{r, echo = F, message = F, warning=F, results = 'hide'}

source("/sfs/qumulo/qhome/sm9dv/dspg20fairfax/src/09_network_descriptives/network_functions.R")
network_stats(vars = "S000", years = 2015, types = types)
data_fairfax <-   get_acs(geography = "tract",
                          variables = "B00001_001",
                          state = "VA",
                          county = 059, 
                          geometry = TRUE)
leaflet_creator <- function(types){
  for(type in types){
  nodelist <- get(paste("nodelist_2015_S000_", type, sep = ""))
  
  fairfax_data <- merge(data_fairfax[, c("GEOID", "geometry")], nodelist[, c("tract", "wtd_deg_cent_out", "wtd_deg_cent_in", "btw_cent", "eigen_cent")], by.x= "GEOID", by.y = "tract", all = TRUE)
  fairfax_data <- st_transform(fairfax_data, 4326)
  
  
  pal_out <- colorQuantile("Blues",5,  domain = fairfax_data$wtd_deg_cent_out)
  pal_in <- colorQuantile("Oranges",5, domain = fairfax_data$wtd_deg_cent_in)
  pal_btw <- colorQuantile("Purples",5,  domain = fairfax_data$btw_cent)
  pal_eigen <- colorQuantile("Greens",5,  domain = fairfax_data$eigen_cent)
  
  
  l_out <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_out(wtd_deg_cent_out), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_out, 
            values =  ~wtd_deg_cent_out,
            title = "Weighted Degree Centrality: Out<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = wtd_deg_cent_out) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
  
  l_in <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_in(wtd_deg_cent_in), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_in, 
            values = ~wtd_deg_cent_in,
            title = "Weighted Degree Centrality: In<br> By Quantile Group", 
            opacity = 1,
            labFormat = function(type = "quantile", cuts = 10, p = wtd_deg_cent_in) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
  
  l_btw <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_btw(btw_cent), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_btw, 
            values = ~btw_cent,
            title = "Betweenness Centrality<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = btw_cent) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
  
  l_eigen <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_eigen(eigen_cent), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_eigen, 
            values = ~eigen_cent,
            title = "Eigen Centrality<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = eigen_cent) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
  
  
  assign(paste("fairfax_2015_S000_", type, sep =  ""), fairfax_data, envir = .GlobalEnv)
  
  assign(paste("pal_in_2015_S000_", type, sep =  ""), pal_in,envir = .GlobalEnv)
  assign(paste("pal_out_2015_S000_", type, sep =  ""), pal_out,envir = .GlobalEnv)
  assign(paste("pal_btw_2015_S000_", type, sep =  ""), pal_btw,envir = .GlobalEnv)
  assign(paste("pal_eigen_2015_S000_", type, sep =  ""), pal_eigen,envir = .GlobalEnv)
  
  assign(paste("l_out_2015_S000_", type, sep =  ""), l_out,envir = .GlobalEnv)
  assign(paste("l_in_2015_S000_", type, sep =  ""), l_in,envir = .GlobalEnv)
  assign(paste("l_btw_2015_S000_", type, sep =  ""), l_btw,envir = .GlobalEnv)
  assign(paste("l_eigen_2015_S000_", type, sep =  ""), l_eigen, envir = .GlobalEnv)
  
}
}
```

### All jobs

```{r, echo = F}
leaflet_creator("all")
fluidRow(column(6, l_out_2015_S000_all), 
         column(6, l_in_2015_S000_all))
fluidRow(column(6, l_btw_2015_S000_all), 
         column(6, l_eigen_2015_S000_all))
```

### Federal Jobs

```{r, echo = F}
type = "federal"
nodelist <- get(paste("nodelist_2015_S000_", type, sep = ""))
  
  
nodelist$color_in <- ifelse(nodelist$wtd_deg_cent_in == 0, "[0]", 
                        ifelse(nodelist$wtd_deg_cent_in < 1000 & 100 <= nodelist$wtd_deg_cent_in, "[100-1000)",
                              ifelse(nodelist$wtd_deg_cent_in < 100 &  10 <= nodelist$wtd_deg_cent_in, "[10-100)", 
                                     "[1-10)")))
nodelist$color_btw <- ifelse(nodelist$btw_cent == 0, "[0]", 
                        ifelse(nodelist$btw_cent < 1000 & 100 <= nodelist$btw_cent, "[100-1000)",
                              ifelse(nodelist$btw_cent < 100 &  10 <= nodelist$btw_cent, "[10-100)", 
                                     "[1-10)")))
nodelist$color_eigen <- ifelse(nodelist$eigen_cent == 0, "[0]", 
                        ifelse(nodelist$eigen_cent < 1 & .1 <= nodelist$eigen_cent, "[.1-1)",
                              ifelse(nodelist$eigen_cent < 0.01 &  0 < nodelist$eigen_cent, "(0-.01)", 
                                     "[.01-.1)")))
  
nodelist$color_eigen <- ordered(nodelist$color_eigen, levels = c("[0]", "(0-.01)","[.01-.1)", "[.1-1)"))
fairfax_data <- merge(data_fairfax[, c("GEOID", "geometry")], nodelist[, c("tract", "wtd_deg_cent_out", "wtd_deg_cent_in", "btw_cent", "eigen_cent", "color_in", "color_btw","color_eigen")], by.x= "GEOID", by.y = "tract", all = TRUE)
  
fairfax_data <- st_transform(fairfax_data, 4326)
pal_out <- colorQuantile("Blues",5,  domain = fairfax_data$wtd_deg_cent_out)
pal_in <- colorFactor(brewer.pal(n = 4, name = "Oranges"), fairfax_data$color_in)
pal_btw <- colorFactor(brewer.pal(n = 4, name = "Purples"), fairfax_data$color_btw)
pal_eigen <- colorFactor(brewer.pal(n = 4, name = "Greens"), fairfax_data$color_eigen)
l_out_2015_S000_federal <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_out(wtd_deg_cent_out), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_out, 
            values =  ~wtd_deg_cent_out,
            title = "Weighted Degree Centrality: Out<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = wtd_deg_cent_out) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
l_in_2015_S000_federal <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_in(color_in), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_in, 
            values = ~color_in,
            title = "Weighted Degree Centrality: In<br>", 
            opacity = 1
          )
l_btw_2015_S000_federal <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_btw(color_btw), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_btw, 
            values = ~color_btw,
            title = "Betweenness Centrality<br>", 
            opacity = 1)
l_eigen_2015_S000_federal <- leaflet(data = fairfax_data, options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_eigen(color_eigen), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_eigen, 
            values = ~color_eigen,
            title = "Eigen Centrality<br>", 
            opacity = 1)
  
data <- read.csv("https://raw.githubusercontent.com/DSPG-Young-Scholars-Program/dspg20STW/sarah/src/edu_knowledge_rothwell/rothwell.csv?token=AMHVUTUDUPE4XEGDLKB5ZZS7DCYH2")

fluidRow(column(6, l_out_2015_S000_federal), 
         column(6, l_in_2015_S000_federal))
fluidRow(column(6, l_btw_2015_S000_federal), 
         column(6, l_eigen_2015_S000_federal))
```

### Private Jobs

```{r, echo = F}
leaflet_creator("private")
fluidRow(column(6, l_out_2015_S000_private), 
         column(6, l_in_2015_S000_private))
fluidRow(column(6, l_btw_2015_S000_private), 
         column(6, l_eigen_2015_S000_private))
```


# References
Citations here.