---
title: 'Fairfax County Labor Markets'
description: 'Symposium Presentation of the 2020 DSPG Project'
weight: 1
output: html_document
---

```{css, echo=FALSE}
/* this chunk of code centers all of the headings */
h1, h2, h3, h4, h5 {
  text-align: center;
}
```


## Fairfax County Labor Markets: Characterizing Local Workforce and Employment Networks


### Our three main goals for the 2020 DSPG Summer Project were to: 

#### (1)  Characterize the Fairfax County workforce, occupation sectors, and industry composition;

#### (2) Identify vulnerable and emerging areas

#### (3) Explore workforce flows


<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

## (1) Characterize the Fairfax County workforce, occupation sectors, and industry composition;

What is the commute type and age of employed Fairfax County residents?

```{r, echo=FALSE, fig.fullwidth=TRUE}
# Maps
knitr::include_app('https://teja.shinyapps.io/app_wrk', height = "650px")
```
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

#### (2a) Identify vulnerable areas

Relative Vulnerability

```{r, warning = F, message = F, echo = F}
library(tidyverse)
library(leaflet)
```

```{r, include = FALSE, echo = FALSE, warning=FALSE, message=FALSE}
# Read in data
vulnerable <- read_rds("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/acs/acs_vuln.Rds") # Sean's ACS 2014/18 vulnerable area indicators data
uiaggregate <- read_rds( "/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/vworks/uiagg_spatial.Rds") # Sean's VWorks 2020 initial unemployment claims spatial data
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
#Summary score leaflet
pal <- colorFactor("Oranges", domain = vulnerable$SummaryScore)

      labels <- lapply(
        paste("<strong>Area: </strong>",
              vulnerable$NAME,
              "<br />",
              "<strong>Workers Unemployed:</strong>",
              round(vulnerable$pct_unempl, 2),
              "<br />",
              "<strong>% Part-Time:</strong>",
              round(vulnerable$pct_part, 2), 
              "<br />",
              "<strong>% High School or Less:</strong>",
              round(vulnerable$pct_hsless, 2), 
              "<br />",
              "<strong>$ Median Worker Income:</strong>",
              formatC(vulnerable$medinc, format = "f", big.mark = ",", digits = 0),
              "<br />",
              "<strong>% In Vulnerable Occupations:</strong>",
              round(vulnerable$pct_occup, 2),
              "<br />",
              "<strong>% In Vulnerable Industries:</strong>",
              round(vulnerable$pct_ind, 2)),
        htmltools::HTML
      )
      
leaflet(data = vulnerable, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(SummaryScore),
              fillOpacity = .8, 
              stroke = FALSE,
              label = labels,
              labelOptions = labelOptions(direction = "bottom",
                                          style = list(
                                            "font-size" = "12px",
                                            "border-color" = "rgba(0,0,0,0.5)",
                                            direction = "auto"))) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~vulnerable$SummaryScore,
            title = "Relative<br>Vulnerability", 
            opacity = 1,
            na.label = "Not Available")
```
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
#### (2b) Identify emerging areas
```{r, include=FALSE}
library(leaflet)
library(dplyr)
library(readxl)
library(readr)
library(sf)
```
```{r, include=FALSE}
# Get data
lodespatial <- read_rds("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/lodes/lodes_spatial.Rds") # Sarah's LODES 2017 workplace + ACS spatial
fortune_geo <- read_sf("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/fortune/fortune_1000/fairfax_fortune_1000_geo.shp") # Fortune 1000

# Get data
#lodespatial <- read_rds("/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/lodes/lodes_spatial.Rds") # Sarah's LODES 2017 workplace + ACS spatial
#fortune_geo <- read_sf("/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/fortune/fortune_1000/fairfax_fortune_1000_geo.shp") # Fortune 1000

employers <- read_excel('/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/vworks/allemployers.xlsx') %>%
#employers <- read_excel('/sfs/qumulo/qhome/tp2sk/Git/dspg20fairfax/data/original/vworks/allemployers.xlsx') %>%
  select_all(~gsub("\\s+|\\.", "_", .)) %>% 
  select_all(tolower) %>% 
  arrange(desc(estimate_of_employees)) %>% 
  slice_head(n = 100) 

```

Where and Who are Fairfax County's Largest Employers?

```{r, warning=FALSE, message=FALSE, echo=FALSE}
pal <- colorQuantile(c("#60999A", "#D9E12B", "#E6CE3A", "#E6A01D", "#E57200"), employers$estimate_of_employees, probs = seq(0, 1, length = 6), right = FALSE)

labels <- lapply(
  paste("<strong>Company name: </strong>",
         employers$business_name,
        "<br />",
        "<strong>6 digit title: </strong>",
        employers$`6_digit_title`,
        "<br />",
        "<strong>Number of employees: </strong>",
        employers$estimate_of_employees),
        htmltools::HTML
  )
leaflet(data = employers, 
        options = leafletOptions(minZoom = 10)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(data = employers, lng = employers$longitude, lat = employers$latitude, 
                   radius = 3.5, color = ~pal(estimate_of_employees), 
                   stroke = FALSE, fillOpacity = 1,
                   label = ~labels, labelOptions = labelOptions(noHide = F, direction = "bottom")) %>%
  addLegend("bottomleft", pal = pal, values = employers$estimate_of_employees, 
            title = "Estimated Number of Employees<br>by Quantile Group", opacity = 1, 
            labFormat = function(type = "quantile", cuts = 5, p = data1$estimate_of_employees) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
                  })
```
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

Do Fortune 1000 Companies Attract Workers?

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# Fortune 1000 HQ plot
pal <- colorQuantile("Oranges", domain = lodespatial$perTotalJobs, probs = seq(0, 1, length = 6), right = FALSE)

leaflet(data = lodespatial, 
        options = leafletOptions(minZoom = 10))%>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(perTotalJobs), fillOpacity = .7, stroke = FALSE) %>%
  addCircleMarkers(data = fortune_geo, popup = fortune_geo$Name, 
                   radius = 3.5,
                   color = "#232D4B",
                   stroke = FALSE, fillOpacity = 1) %>%
  addLegend("bottomleft", 
            pal = pal, 
            values =  ~perTotalJobs,
            title = "Percent of All<br>Fairfax County Jobs,<br>by Quintile Group", 
            opacity = 1,
            labFormat = function(type = "quantile", cuts = 5, p = perTotalJobs) {
                    n = length(cuts)
                    paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
#### (3a) Networks: Examining Job Flows From Outside of Fairfax County

Where are the Major Counties <br/> Outside of Fairfax that Workers Reside In? 
```{r warning = F, message = F, results = 'hide', echo = F}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(forcats)
library(sf)
library(tmap)
library(leaflet)
library(stplanr)
library(tigris)
```

```{r warning = F, message = F, results = 'hide', echo = F}
# Read in data
#jobs_all <- read_csv("~/git/dspg20fairfax/data/od/jobsall.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))
jobs_all <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/od/jobsall.csv", 
                     col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))

#va_xw <- read_csv("~/git/dspg20fairfax/data/original/va_xwalk.csv")
va_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/va_xwalk.csv")
```

```{r warning = F, message = F, echo = F} 
# Parsing GEOID
jobs_all$h_state <- substr(jobs_all$h_geocode_tract, 1, 2)
jobs_all$h_county <- substr(jobs_all$h_geocode_tract, 3, 5)
jobs_all$h_tract <- substr(jobs_all$h_geocode_tract, 6, 11)
jobs_all$h_stcty <- substr(jobs_all$h_geocode_tract, 1, 5)

# Now parsing workplace GEOIDs
jobs_all$w_state <- substr(jobs_all$w_geocode_tract, 1, 2)
jobs_all$w_county <- substr(jobs_all$w_geocode_tract, 3, 5)
jobs_all$w_tract <- substr(jobs_all$w_geocode_tract, 6, 11)
jobs_all$w_stcty <- substr(jobs_all$w_geocode_tract, 1, 5)
```

```{r warning = F, message = F, results = 'hide', echo = F}
nc_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/nc_xwalk.csv", col_types = cols(cty = col_character()))
dc_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/dc_xwalk.csv", col_types = cols(cty = col_character()))
pa_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/pa_xwalk.csv", col_types = cols(cty = col_character()))
wv_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/wv_xwalk.csv", col_types = cols(cty = col_character()))
md_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/md_xwalk.csv", col_types = cols(cty = col_character()))
oh_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/oh_xwalk.csv", col_types = cols(cty = col_character()))
ny_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/ny_xwalk.csv", col_types = cols(cty = col_character()))
fl_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/fl_xwalk.csv", col_types = cols(cty = col_character()))
nj_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/nj_xwalk.csv", col_types = cols(cty = col_character()))
sc_xw <- read_csv("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/data/original/xwalk/sc_xwalk.csv", col_types = cols(cty = col_character()))
```

```{r warning = F, message = F, results = 'hide', echo = F}
# Loading top 10 states to map flows
va_points <- tracts("VA", class = "sf")
md_points <- tracts("MD", class = "sf")
dc_points <- tracts("DC", class = "sf")
wv_points <- tracts("WV", class = "sf")
pa_points <- tracts("PA", class = "sf")
nc_points <- tracts("NC", class = "sf")
nj_points <- tracts("NJ", class = "sf")
ny_points <- tracts("NY", class = "sf")
oh_points <- tracts("OH", class = "sf")
sc_points <- tracts("SC", class = "sf")
fl_points <- tracts("FL", class = "sf")
```

```{r warning = F, message = F, results = 'hide', echo = F}
#this means all jobs where residence is outside of Fairfax but workplace is in Fairfax
flows_in <- jobs_all %>% filter(h_stcty != "51059" & w_stcty == "51059")
bool_cond <- jobs_all$h_stcty != "51059" & jobs_all$w_stcty == "51059"

percents <- data.frame("type" = c("From outside Fairfax County", "From within Fairfax County"), "shares" = c((mean(bool_cond) * 100), ((1 - mean(bool_cond)) * 100)))

#NC done
essential <- nc_xw %>%
  select(cty, ctyname)
nc_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#dc done
essential <- dc_xw %>%
  select(cty, ctyname)
dc_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#pa done
essential <- pa_xw %>%
  select(cty, ctyname)
pa_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#wv done
essential <- wv_xw %>%
  select(cty, ctyname)
wv_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#md done
essential <- md_xw %>%
  select(cty, ctyname)
md_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#oh done
essential <- oh_xw %>%
  select(cty, ctyname)
oh_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#ny done
essential <- ny_xw %>%
  select(cty, ctyname)
ny_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#fl done
essential <- fl_xw %>%
  select(cty, ctyname)
fl_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#nj done
essential <- nj_xw %>%
  select(cty, ctyname)
nj_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#sc done
essential <- sc_xw %>%
  select(cty, ctyname)
sc_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
top_ten_with_names <- rbind(nc_names, dc_names, pa_names, wv_names, md_names, oh_names, ny_names, fl_names, nj_names, sc_names)
agg_cnty_other_with_names <- top_ten_with_names %>%
  group_by(ctyname) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup()
agg_cnty_other_trimmed <- agg_cnty_other_with_names[order(-agg_cnty_other_with_names$S000), ][1:10, ]
agg_cnty_other_trimmed %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "#0E879C") +
  ggtitle("Number of Jobs Flowing into Fairfax from\nOther Counties in Other States") +
  theme_minimal() +
  xlab("Number of Jobs") + 
  ylab("County Name") +
  scale_x_continuous(labels = scales::comma) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) 
```
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
What jobs come from other states?
```{r, echo=FALSE, warning=FALSE, message=FALSE, include =FALSE}
flows_in_state <- flows_in[flows_in$h_state != "51", ] %>%
  group_by(h_state, w_state) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup() %>%
  arrange(-S000)
#get states names
flows_in_state <- flows_in_state[1:10, ]
flows_in_state$State <- c("Maryland", "District of Columbia", "West Virginia", "Pennsylvania", "North Carolina", "New Jersey", "New York", "Ohio", "South Carolina", "Florida")

states_flow_lines <-st_as_sf(od2line(flows_in_state, states()[3:10]))
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
tmap_mode("view")
pal <- viridis::viridis(n = 3, option = "C")
top10states_flowsmap <- tm_shape(states_flow_lines %>% select(State, w_state, S000, geometry)) +
  tm_lines(lwd = 3.5, 
           col = "S000", 
           title.col = "Number of Jobs", 
           style = "quantile",
           palette = c("#E6A01D", "#E57200"),
           popup.vars = c("Number of Jobs" = "S000")) +
  tm_basemap("OpenStreetMap.HOT") +
  tm_view(set.zoom.limits = c(5, 18), basemaps.alpha = 0.5)
top10states_flowsmap
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
#### Using Network Analysis to Study Fairfax County Job Flows

All Job Flows in Fairfax County

```{r, include=FALSE}
library(tidycensus)
library(leaflet)
library(sf)
library(shiny)
library(fabricatr)
library(RColorBrewer)

#source("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/src/09_network_descriptives/network_functions.R")
source("/sfs/qumulo/qhome/kb7hp/git/dspg20fairfax/src/09_network_descriptives/network_functions.R")
network_stats(vars = "S000", years = 2015, types = types)
```

```{r, echo = F}
leaflet_creator("all")
fluidRow(column(6, l_out_2015_S000_all), 
         column(6, l_in_2015_S000_all))
fluidRow(column(6, l_btw_2015_S000_all), 
         column(6, l_eigen_2015_S000_all))
```

