---
title: "Network Maps"
author: "Sarah McDonald"
date: "7/10/2020"
output: html_document
---

### Interpretation of Centrality Measures:

_Weighted Out-Degree Centrality:_ These plots highlight census tracts from which individuals depart to get to their jobs in other tracts.

_Weighted In-Degree Centrality:_ These plots highlight census tracts that are work destinations.

_Betweenness Centrality:_ Betweenness centrality measures if a particular tract lies on paths between other tracts. 

_Eigenvector Centrality:_ Also known as rank centrality, measures importance of a tract based on if neighboring connections are important, in other words, have lots of connections.

```{r, echo = F, message = F, warning=F, results = 'hide'}
library(tidycensus)
library(leaflet)
library(sf)
library(shiny)
library(fabricatr)
library(RColorBrewer)

source("/sfs/qumulo/qhome/sm9dv/dspg20fairfax/src/09_network_descriptives/network_functions.R")

network_stats(vars = "S000", years = 2015, types = types)

data_fairfax <-   get_acs(geography = "tract",
                          variables = "B00001_001",
                          state = "VA",
                          county = 059, 
                          geometry = TRUE)

leaflet_creator <- function(types){
  for(type in types){
  nodelist <- get(paste("nodelist_2015_S000_", type, sep = ""))
  
  fairfax_data <- merge(data_fairfax[, c("GEOID", "geometry")], nodelist[, c("tract", "wtd_deg_cent_out", "wtd_deg_cent_in", "btw_cent", "eigen_cent")], by.x= "GEOID", by.y = "tract", all = TRUE)
  fairfax_data <- st_transform(fairfax_data, 4326)
  
  
  pal_out <- colorQuantile("Blues",5,  domain = fairfax_data$wtd_deg_cent_out)

  pal_in <- colorQuantile("Oranges",5, domain = fairfax_data$wtd_deg_cent_in)

  pal_btw <- colorQuantile("Purples",5,  domain = fairfax_data$btw_cent)

  pal_eigen <- colorQuantile("Greens",5,  domain = fairfax_data$eigen_cent)
  
  
  l_out <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_out(wtd_deg_cent_out), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_out, 
            values =  ~wtd_deg_cent_out,
            title = "Weighted Degree Centrality: Out<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = wtd_deg_cent_out) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
  
  l_in <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_in(wtd_deg_cent_in), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_in, 
            values = ~wtd_deg_cent_in,
            title = "Weighted Degree Centrality: In<br> By Quantile Group", 
            opacity = 1,
            labFormat = function(type = "quantile", cuts = 10, p = wtd_deg_cent_in) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })

  
  l_btw <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_btw(btw_cent), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_btw, 
            values = ~btw_cent,
            title = "Betweenness Centrality<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = btw_cent) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })

  
  l_eigen <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_eigen(eigen_cent), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_eigen, 
            values = ~eigen_cent,
            title = "Eigen Centrality<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = eigen_cent) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })
  
  
  assign(paste("fairfax_2015_S000_", type, sep =  ""), fairfax_data, envir = .GlobalEnv)
  
  assign(paste("pal_in_2015_S000_", type, sep =  ""), pal_in,envir = .GlobalEnv)
  assign(paste("pal_out_2015_S000_", type, sep =  ""), pal_out,envir = .GlobalEnv)
  assign(paste("pal_btw_2015_S000_", type, sep =  ""), pal_btw,envir = .GlobalEnv)
  assign(paste("pal_eigen_2015_S000_", type, sep =  ""), pal_eigen,envir = .GlobalEnv)
  
  assign(paste("l_out_2015_S000_", type, sep =  ""), l_out,envir = .GlobalEnv)
  assign(paste("l_in_2015_S000_", type, sep =  ""), l_in,envir = .GlobalEnv)
  assign(paste("l_btw_2015_S000_", type, sep =  ""), l_btw,envir = .GlobalEnv)
  assign(paste("l_eigen_2015_S000_", type, sep =  ""), l_eigen, envir = .GlobalEnv)
  
}
}



```

### All jobs

```{r, echo = F}

leaflet_creator("all")


fluidRow(column(6, l_out_2015_S000_all), 
         column(6, l_in_2015_S000_all))
fluidRow(column(6, l_btw_2015_S000_all), 
         column(6, l_eigen_2015_S000_all))
```

### Federal Jobs

```{r, echo = F}

type = "federal"

nodelist <- get(paste("nodelist_2015_S000_", type, sep = ""))
  
  
nodelist$color_in <- ifelse(nodelist$wtd_deg_cent_in == 0, "[0]", 
                        ifelse(nodelist$wtd_deg_cent_in < 1000 & 100 <= nodelist$wtd_deg_cent_in, "[100-1000)",
                              ifelse(nodelist$wtd_deg_cent_in < 100 &  10 <= nodelist$wtd_deg_cent_in, "[10-100)", 
                                     "[1-10)")))
nodelist$color_btw <- ifelse(nodelist$btw_cent == 0, "[0]", 
                        ifelse(nodelist$btw_cent < 1000 & 100 <= nodelist$btw_cent, "[100-1000)",
                              ifelse(nodelist$btw_cent < 100 &  10 <= nodelist$btw_cent, "[10-100)", 
                                     "[1-10)")))
nodelist$color_eigen <- ifelse(nodelist$eigen_cent == 0, "[0]", 
                        ifelse(nodelist$eigen_cent < 1 & .1 <= nodelist$eigen_cent, "[.1-1)",
                              ifelse(nodelist$eigen_cent < 0.01 &  0 < nodelist$eigen_cent, "(0-.01)", 
                                     "[.01-.1)")))
  
nodelist$color_eigen <- ordered(nodelist$color_eigen, levels = c("[0]", "(0-.01)","[.01-.1)", "[.1-1)"))

fairfax_data <- merge(data_fairfax[, c("GEOID", "geometry")], nodelist[, c("tract", "wtd_deg_cent_out", "wtd_deg_cent_in", "btw_cent", "eigen_cent", "color_in", "color_btw","color_eigen")], by.x= "GEOID", by.y = "tract", all = TRUE)
  
fairfax_data <- st_transform(fairfax_data, 4326)


pal_out <- colorQuantile("Blues",5,  domain = fairfax_data$wtd_deg_cent_out)
pal_in <- colorFactor(brewer.pal(n = 4, name = "teal"), fairfax_data$color_in)
pal_btw <- colorFactor(brewer.pal(n = 4, name = "Purples"), fairfax_data$color_btw)
pal_eigen <- colorFactor(brewer.pal(n = 4, name = "Greens"), fairfax_data$color_eigen)


l_out_2015_S000_federal <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_out(wtd_deg_cent_out), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_out, 
            values =  ~wtd_deg_cent_out,
            title = "Weighted Degree Centrality: Out<br> By Quantile Group", 
            opacity = 1, 
            labFormat = function(type = "quantile", cuts = 10, p = wtd_deg_cent_out) {
              n = length(cuts)
              paste0("[", round(cuts[-n], 3), " &ndash; ", round(cuts[-1], 3), ")")
            })

l_in_2015_S000_federal <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_in(color_in), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_in, 
            values = ~color_in,
            title = "Weighted Degree Centrality: In<br>", 
            opacity = 1
          )


l_btw_2015_S000_federal <- leaflet(data = fairfax_data,options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_btw(color_btw), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_btw, 
            values = ~color_btw,
            title = "Betweenness Centrality<br>", 
            opacity = 1)


l_eigen_2015_S000_federal <- leaflet(data = fairfax_data, options = leafletOptions(minZoom = 10))%>%
  addTiles("https://{s}.basemaps.cartocdn.com/rastertiles/voyager_labels_under/{z}/{x}/{y}.png") %>%
  addPolygons(fillColor = ~pal_eigen(color_eigen), fillOpacity = .8, stroke = FALSE) %>%
  addLegend("bottomleft", 
            pal = pal_eigen, 
            values = ~color_eigen,
            title = "Eigen Centrality<br>", 
            opacity = 1)
  

fluidRow(column(6, l_out_2015_S000_federal), 
         column(6, l_in_2015_S000_federal))
fluidRow(column(6, l_btw_2015_S000_federal), 
         column(6, l_eigen_2015_S000_federal))
```

### Private Jobs

```{r, echo = F}
leaflet_creator("private")

fluidRow(column(6, l_out_2015_S000_private), 
         column(6, l_in_2015_S000_private))
fluidRow(column(6, l_btw_2015_S000_private), 
         column(6, l_eigen_2015_S000_private))
```

