---
title: "09_network_descriptives_intra_fairfax"
author: "Sarah McDonald"
date: "6/29/2020"
#output: html_document
output: md_document
always_allow_html: true
---

Load packages
```{r, warning=F, message=F}
library(dplyr)
library(igraph)
library(lubridate)
```

Load data frame and filter for residence AND workplace within Fairfax County. 

```{r}
data <- read.csv("/sfs/qumulo/qhome/sm9dv/dspg20fairfax/data/od/jobsall.csv")

# focus on within county first
data <- data %>% 
  filter(grepl("^51059.+", w_geocode_tract) & grepl("^51059.+", h_geocode_tract))
```

# Edge List

Create edge list, which defines the "from" and "to" relationships of the Fairfax workforce. We define "from" as residence and "to" as workplace. For this script, we will define the weight as S000, which is the total number of jobs, in other words,the total number of jobs for which a worker lives in the "from" Census tract and commutes to the workplace in the "to" Census tract (check interpretation). 

```{r, warning=F, message =F}
edgelist <- data %>% 
  select(h_geocode_tract, w_geocode_tract, S000) %>% 
  rename(from = h_geocode_tract, to = w_geocode_tract, weight = S000) %>% 
  group_by(from, to) %>% 
  summarize(weight = sum(weight)) %>% 
  arrange(-weight)
```

Next, we create the __igraph__ network object. Simple graphs are graphs which do not contain loop and multiple edges. 
The function below keeps the loops, which would indicate working and living in the same tract, but removes multiple edges, because we will start wth a simple, directed graph.


```{r}
network <- simplify(graph.data.frame(edgelist, directed = TRUE), 
                         remove.loops = FALSE, 
                         edge.attr.comb = igraph_opt("edge.attr.comb"))
is_weighted(network)
is_directed(network)


#D: directed
#N: Named
#W: Weighted

print(network)
```


### Summary Statistics

Next, we will work through summary statistics of our network with descriptions of each, in the context of the origin-destination data.

```{r}
network_stats <- data.frame(var="S000") 
```

##### Counts
The _order_ of a graph indicates the number of nodes (or vertices). In this case, the nodes are Census tracts.

```{r}
network_stats$node_count <- gorder(network)  
print(network_stats$node_count )
```

The _size_ of graph indicates the number of edges (or links). Relevant to OD data, the an edge represents a residence-workplace relationship between Census tracts.

```{r}
network_stats$edge_count <- gsize(network)  
print(network_stats$edge_count)
```

We can also add up the sum of weights for a count of jobs represented in the graph.

```{r}
network_stats$jobs_count <- sum(edgelist$weight)
print(network_stats$jobs_count)
```

The __degree__ indicates the number of adjacent edges. An adjacent edge shares a common vertex. In relation to the OD, the degree function produces a count of links for each node, in this case, for each Census tract. The average degree is printed.
```{r}
network_stats$mean_deg_total <- mean(igraph::degree(network))
network_stats$mean_deg_out <- mean(igraph::degree(network, mode = "out"))
network_stats$mean_deg_in <- mean(igraph::degree(network, mode = "in"))

#mean(sna::degree(intergraph::asNetwork(network), cmode="freeman"))
#mean(sna::degree(intergraph::asNetwork(network), cmode="indegree"))
#mean(sna::degree(intergraph::asNetwork(network), cmode="outdegree"))

knitr::kable(network_stats[, c("mean_deg_total", "mean_deg_out", "mean_deg_in")])

hist(igraph::degree(network), breaks = seq(0,425, 25), main="Histogram of Node Degree")
```

__Betweenness__ is a measure of centrality, which counts the number of shortest paths (geodesic) going through a node. 

```{r}
network_stats$mean_btw <- mean(round(sna::betweenness(intergraph::asNetwork(network), cmode="directed"), 4))

print(network_stats$mean_btw)

hist(round(sna::betweenness(intergraph::asNetwork(network), cmode="directed"), 4), main = "Histogram of Betweenness", breaks = seq(0, 500, 25))
```

##### Isolates, Dyads, and Triads 

Isolates are non-connected nodes, which means either the workplace or residence did not have a corresponding relationship.
```{r}
network_stats$isolates <- sum(igraph::degree(simplify(network))==0)
```

Triads
```{r}
network_stats$triads_003 <- igraph::triad.census(network)[1] # empty graph, no connections
network_stats$triads_012 <- igraph::triad.census(network)[2]
network_stats$triads_102 <- igraph::triad.census(network)[3] #graph with a mutual connection between two vertices.
network_stats$triads_021D <- igraph::triad.census(network)[4]
network_stats$triads_021U <- igraph::triad.census(network)[5]
network_stats$triads_021C <- igraph::triad.census(network)[6]
network_stats$triads_111D <- igraph::triad.census(network)[7]
network_stats$triads_111U <- igraph::triad.census(network)[8]
network_stats$triads_030T <- igraph::triad.census(network)[9]
network_stats$triads_030C <- igraph::triad.census(network)[10]
network_stats$triads_201 <- igraph::triad.census(network)[11] #graph with two mutual connection between two vertices.
network_stats$triads_120D <- igraph::triad.census(network)[12]
network_stats$triads_120U <- igraph::triad.census(network)[13]
network_stats$triads_120C <- igraph::triad.census(network)[14]
network_stats$triads_210 <- igraph::triad.census(network)[15]
network_stats$triads_300 <- igraph::triad.census(network)[16] # complete graph

knitr::kable(network_stats[, c(colnames(network_stats)[grepl("^triads_", colnames(network_stats))])][1:8])
knitr::kable(network_stats[, c(colnames(network_stats)[grepl("^triads_", colnames(network_stats))])][9:16])
```

Dyads 

Mutual Dyads are the count of pairs with mutual connections.
Asymmetric Dyads are the count of pairs with non-mutual connections.
Null Dyads are the count of pairs with no connections.

```{r}
network_stats$dyad_mut <- igraph::dyad_census(network)$mut
network_stats$dyad_asym <- igraph::dyad_census(network)$asym
network_stats$dyad_null <- igraph::dyad_census(network)$null

knitr::kable(network_stats[, c(colnames(network_stats)[grepl("^dyad_", colnames(network_stats))])])
```


##### Density and Transitivity

The __diameter__ of a graph is the length of the longest shortest path (geodesic) passing through a node.  
```{r}
network_stats$diameter <- diameter(network,directed=TRUE, 
                                   unconnected=if (network_stats$isolates == 0) {FALSE} else {TRUE})
```


Mean Distance: The average length of all the shortest paths from or to the nodes in the network.
```{r}
network_stats$mean_distance <- mean_distance(network, directed = TRUE, 
                                   unconnected = if (network_stats$isolates == 0) {FALSE} else {TRUE})

```


Edge Density: ratio of the number of edges and the number of possible edges. Density ranges from 0 (minimum) to 1 (maximum).
```{r}
network_stats$density <- edge_density(network, loops=TRUE) 
print(network_stats$density)
```

Transitivity: The probability that the adjacent vertices of a vertex are connected
```{r}
network_stats$transitivity_global <- transitivity(network, weights = TRUE, type = "global")
network_stats$transitivity_local_avg <- mean(transitivity(network, weights = TRUE, type = "local"))
#network_stats$transitivity_barrat <- transitivity(network, weights = TRUE, type = "weighted") # weighted (this does not work)

knitr::kable(network_stats[, c("transitivity_global", "transitivity_local_avg")])
```


### Community Detection
Louvain Method: The multi-level modularity optimization algorithm 
```{r}
# this doesn't work for directed networks
louvain <- igraph::cluster_louvain(as.undirected(network))
network_stats$louvain <- modularity(louvain)
network_stats$louvain_scaled <- modularity(louvain) / gorder(network)
network_stats$louvain_logged <- modularity(louvain) / log(gorder(network))

knitr::kable(network_stats[, c("louvain","louvain_scaled", "louvain_logged")])
```

Fast and Greedy: searches for a dense sub-graph

```{r}
# does not work for directed graphs
fstgrdy <- cluster_fast_greedy(as.undirected(network))
network_stats$fstgrdy <- modularity(fstgrdy)
network_stats$fstgrdy_scaled <- modularity(fstgrdy) / gorder(network)
network_stats$fstgrdy_logged <- modularity(fstgrdy) / log(gorder(network))

knitr::kable(network_stats[, c("fstgrdy","fstgrdy_scaled", "fstgrdy_logged")])
```

### Centrality

Degree Centrality: Centralize a graph according to the degrees of vertices
Closenses Centrality: Centralize a graph according to the closeness of vertices
Betweenness Centrality: Centralize a graph according to the betweenness of vertices
Eigenvector Centrality: Centralize a graph according to the eigenvector centrality of vertices

```{r}
network_stats$centr_deg_total <- round(centr_degree(network, mode = "all")$centralization, 3)
network_stats$centr_deg_out <- round(centr_degree(network, mode = "out")$centralization, 3)
network_stats$centr_deg_in <- round(centr_degree(network, mode = "in")$centralization, 3)

network_stats$centr_clo_total <- round(centr_clo(network, mode = "all")$centralization, 3)
network_stats$centr_clo_out <- round(centr_clo(network, mode = "out")$centralization, 3)
network_stats$centr_clo_in <- round(centr_clo(network, mode = "in")$centralization, 3)

network_stats$centr_betw_total <- round(centr_betw(network, directed = TRUE)$centralization, 3)

network_stats$centr_eigen_total <- round(centr_eigen(network, directed = TRUE)$centralization, 3)

```

### Decomposition Statistics
```{r}
decomposition_stats <- table(sapply(decompose.graph(network), vcount))
```


# Nodelist

```{r}
nodelist <- data.frame(id = c(1:(igraph::vcount(network))), tract = igraph::V(network)$name)
```

### degree, weighted degree, k core and modularity 


Degree Centrality and Weighted Degree Centrality (All/Out/In):
```{r}
nodelist$deg_cent_total <- igraph::degree(network, mode = "all")
nodelist$deg_cent_out <- igraph::degree(network, mode = "out")
nodelist$deg_cent_in <- igraph::degree(network, mode = "in")

nodelist$wtd_deg_cent_total <- igraph::strength(network, mode = "all")
nodelist$wtd_deg_cent_out <- igraph::strength(network, mode = "out")
nodelist$wtd_deg_cent_in <- igraph::strength(network, mode = "in")

```

Betweeness Centrality:
```{r}
nodelist$btw_cent <- round(sna::betweenness(intergraph::asNetwork(network), cmode="directed"), 4)
```

Closeness Centrality (All/In/Out): 
```{r}
nodelist$close_cent_total <- igraph::centr_clo(network, mode = "all")$res 
nodelist$close_cent_out <- igraph::centr_clo(network, mode = "out")$res  
nodelist$close_cent_in <- igraph::centr_clo(network, mode = "in")$res 
```

Eigenvector Centrality: 
```{r}
nodelist$eigen_cent <- igraph::eigen_centrality(network, directed = TRUE)$vector
```

### Membership

```{r}
components <- components(network)
fstgrdy <- fastgreedy.community(as.undirected(network))
louvain <- cluster_louvain(as.undirected(network))

nodelist$component <- components$membership
nodelist$louvain_comm <- louvain$membership
nodelist$fstgrdy_comm <- fstgrdy$membership
```

## Other Node-Level Measures 

```{r}
nodelist$page_rank <- page_rank(network, directed=TRUE, weights=NULL)$vector
nodelist$alpha_cent <- alpha_centrality(network, weights=NULL)
nodelist$power_cent <- power_centrality(network, rescale=TRUE)
nodelist$subgraph_cent <- subgraph_centrality(network,diag=TRUE)
nodelist$auth_score <- authority_score(network, scale = TRUE, weights = NULL)$vector
nodelist$hub_score <- hub_score(network, scale = TRUE, weights = NULL)$vector
nodelist$local_trans <- transitivity(network, type = "local", weights = NULL)
nodelist$in_ecc_cent <- eccentricity(network, mode = "in")
nodelist$out_ecc_cent <- eccentricity(network, mode = "out")
nodelist$load_cent <- sna::loadcent(get.adjacency(network,sparse=F),
                                    gmode="digraph",diag=TRUE,cmode="directed",rescale=TRUE)
nodelist$info_cent <- sna::infocent(get.adjacency(network,sparse=F),
                                    gmode="digraph",diag=TRUE,cmode="directed",rescale=TRUE)
nodelist$stress_cent <- sna::stresscent(get.adjacency(network,sparse=F),
                                        gmode="digraph",diag=TRUE,cmode="directed",rescale=TRUE)
nodelist$gilschmidt <- sna::gilschmidt(get.adjacency(network,sparse=F),
                                       gmode = "digraph", diag = TRUE, normalize=TRUE)
nodelist$hyper_index <- netrankr::hyperbolic_index(network)
nodelist$comm_bet <- centiserve::communibet(network, normalized = TRUE)
nodelist$decay_in <- centiserve::decay(network, mode="in", weights=NULL)
nodelist$decay_out <- centiserve::decay(network, mode="out", weights=NULL)
nodelist$diffus_indeg <- centiserve::diffusion.degree(network,mode="in",loops=TRUE)
nodelist$diffus_outdeg <- centiserve::diffusion.degree(network,mode="out",loops=TRUE)     
nodelist$entropy_in <- 1/centiserve::entropy(network,mode="in",weights=NULL)
nodelist$entropy_out <- 1/centiserve::entropy(network,mode="out",weights=NULL)
nodelist$geokpath_in <- centiserve::geokpath(network,mode="in",weights=NULL)
nodelist$geokpath_out <- centiserve::geokpath(network,mode="out",weights=NULL)
nodelist$laplacian_in <- centiserve::laplacian(network,mode="in",loops=TRUE)
nodelist$laplacian_out <- centiserve::laplacian(network,mode="out",loops=TRUE)
nodelist$leverage_in <- centiserve::leverage(network,mode="in",loops=TRUE)  
nodelist$leverage_out <- centiserve::leverage(network,mode="out",loops=TRUE)  
nodelist$lin_cent_in <- centiserve::lincent(network,mode="in",weights=NULL)
nodelist$lin_cent_out <- centiserve::lincent(network,mode="out",weights=NULL)
nodelist$lobby_in <- centiserve::lobby(network,mode="in",loops=TRUE) 
nodelist$lobby_out <- centiserve::lobby(network,mode="out",loops=TRUE) 
nodelist$markov_cent <- centiserve::markovcent(network) 
nodelist$bottleneck_in <- centiserve::bottleneck(network, mode="in")
nodelist$bottleneck_out <- centiserve::bottleneck(network, mode="out")
nodelist$radiality_in <- centiserve::radiality(network, mode="in",weights=NULL)   
nodelist$radiality_out <- centiserve::radiality(network, mode="out",weights=NULL)
nodelist$k_core_in <- coreness(network, mode="in")
nodelist$k_core_out <- coreness(network, mode="out")
nodelist$mnc_in <- centiserve::mnc(network, mode="in")
nodelist$mnc_out <- centiserve::mnc(network, mode="out")
nodelist$dmnc_in <- centiserve::dmnc(network, mode="in")
nodelist$dmnc_out <- centiserve::dmnc(network, mode="out")
nodelist$epc <- centiserve::epc(network) 
nodelist$topocoefficient <- 1/centiserve::topocoefficient(network)
```


### Sources
http://www.people.vcu.edu/~gasmerom/MAT131/graphs.html#:~:text=Two%20vertices%20are%20said%20to,edge%20(arc)%20connecting%20them.&text=Adjacent%20edges%20are%20edges%20that%20share%20a%20common%20vertex.&text=The%20degree%20of%20a%20vertex,edges%20incident%20with%20that%20vertex.&text=A%20graph%20is%20said%20to,are%20joined%20by%20a%20path.

https://github.com/uva-bi-sdad/oss-2020/blob/master/src/github-network-analysis/02_international-collaboration/03_country-networks/03_ctry-nets-cum-analysis.Rmd

https://www.jessesadler.com/post/network-analysis-with-r/

https://kateto.net/wp-content/uploads/2018/03/R%20for%20Networks%20Workshop%20-%20Ognyanova%20-%202018.pdf

https://kateto.net/network-visualization

https://ebookcentral-proquest-com.proxy01.its.virginia.edu/lib/uva/reader.action?docID=4199226&ppg=185

https://lehd.ces.census.gov/data/lodes/LODES7/LODESTechDoc7.4.pdf

https://igraph.org/r/doc/igraph.pdf












