---
title: "09_network_descriptives_intra_fairfax"
author: "Sarah McDonald"
date: "6/29/2020"
output: html_document
#output: md_document
#always_allow_html: true
---

Load packages
```{r, warning=F, message=F}
library(dplyr)
library(igraph)
library(lubridate)
library(sna)
library(intergraph)
```

Load data frame and filter for residence AND workplace within Fairfax County. 

```{r}
data <- read.csv("/sfs/qumulo/qhome/sm9dv/dspg20fairfax/data/od/jobsall.csv")

# focus on within county first
data <- data %>% 
  filter(grepl("^51059.+", w_geocode_tract) & grepl("^51059.+", h_geocode_tract))
```


Create edge list, which defines the "from" and "to" relationships of the Fairfax workforce. We define "from" as residence and "to" as workplace. For this script, we will define the weight as S000, which is the total number of jobs, in other words,the total number of jobs for which a worker lives in the "from" Census tract and commutes to the workplace in the "to" Census tract (check interpretation). 

```{r, warning=F, message =F}
edgelist <- data %>% 
  select(h_geocode_tract, w_geocode_tract, S000) %>% 
  rename(from = h_geocode_tract, to = w_geocode_tract, weight = S000) %>% 
  group_by(from, to) %>% 
  summarize(weight = sum(weight)) %>% 
  arrange(-weight)
```

Next, we create the __igraph__ network object. Simple graphs are graphs which do not contain loop and multiple edges. 
The function below keeps the loops, which would indicate working and living in the same tract, but removes multiple edges, because we will start wth a simple, directed graph.


```{r}
network <- simplify(graph.data.frame(edgelist, directed = TRUE), 
                         remove.loops = FALSE, 
                         edge.attr.comb = igraph_opt("edge.attr.comb"))
#D: directed
#N: Named
#W: Weighted

print(network)
```


### Summary Statistics

Next, we will work through summary statistics of our network with descriptions of each, in the context of the origin-destination data.

```{r}
network_stats <- data.frame(var="S000") 
```

##### Counts
The _order_ of a graph indicates the number of nodes (or vertices). In this case, the nodes are Census tracts.

```{r}
network_stats$node_count <- gorder(network)  
print(network_stats$node_count )
```

The _size_ of graph indicates the number of edges (or links). Relevant to OD data, the an edge represents a residence-workplace relationship between Census tracts.

```{r}
network_stats$edge_count <- gsize(network)  
print(network_stats$edge_count)
```

We can also add up the sum of weights for a count of jobs represented in the graph.

```{r}
network_stats$jobs_count <- sum(edgelist$weight)
print(network_stats$jobs_count)
```

The __degree__ indicates the number of adjacent edges. An adjacent edge shares a common vertex. In relation to the OD, the degree function produces a count of links for each node, in this case, for each Census tract. The average degree is printed.
```{r}
network_stats$mean_deg_total <- mean(igraph::degree(network))
network_stats$mean_deg_out <- mean(igraph::degree(network, mode = "out"))
network_stats$mean_deg_in <- mean(igraph::degree(network, mode = "in"))
print(network_stats$mean_deg_total)

hist(igraph::degree(network), breaks = seq(0,425, 25), main="Histogram of Node Degree")
```

__Betweenness__ is a measure of centrality, which counts the number of shortest paths (geodesic) going through a node. 

```{r}
network_stats$mean_btw <- mean(round(sna::betweenness(intergraph::asNetwork(network), cmode="directed"), 4))
print(network_stats$mean_btw)
hist(round(sna::betweenness(intergraph::asNetwork(network), cmode="directed"), 4), main = "Histogram of Betweenness", breaks = seq(0, 500, 25))
```

##### Isolates, Dyads, and Triads 

Isolates are non-connected nodes, which means either the workplace or residence did not have a corresponding relationship.
```{r}
network_stats$isolates <- sum(igraph::degree(simplify(network))==0)
```

Triads
Why do we care about triads?

```{r}
network_stats$triads_003 <- igraph::triad.census(network)[1] # empty graph, no connections
network_stats$triads_012 <- igraph::triad.census(network)[2]
network_stats$triads_102 <- igraph::triad.census(network)[3] #graph with a mutual connection between two vertices.
network_stats$triads_021D <- igraph::triad.census(network)[4]
network_stats$triads_021U <- igraph::triad.census(network)[5]
network_stats$triads_021C <- igraph::triad.census(network)[6]
network_stats$triads_111D <- igraph::triad.census(network)[7]
network_stats$triads_111U <- igraph::triad.census(network)[8]
network_stats$triads_030T <- igraph::triad.census(network)[9]
network_stats$triads_030C <- igraph::triad.census(network)[10]
network_stats$triads_201 <- igraph::triad.census(network)[11] #graph with two mutual connection between two vertices.
network_stats$triads_120D <- igraph::triad.census(network)[12]
network_stats$triads_120U <- igraph::triad.census(network)[13]
network_stats$triads_120C <- igraph::triad.census(network)[14]
network_stats$triads_210 <- igraph::triad.census(network)[15]
network_stats$triads_300 <- igraph::triad.census(network)[16] # complete graph
```



Dyads (this will change with directionality)

```{r}
# The number of pairs with mutual connections.
network_stats$dyad_mut <- igraph::dyad_census(network)$mut

# The number of pairs with non-mutual connections.
network_stats$dyad_asym <- igraph::dyad_census(network)$asym

# The number of pairs with no connections.
network_stats$dyad_null <- igraph::dyad_census(network)$null
```


##### Density and Transitivity

The __diameter__ of a graph is the length of the longest shortest path (geodesic) passing through a node.  
```{r}
network_stats$diameter <- diameter(network,directed=TRUE, 
                                   unconnected=if (network_stats$isolates == 0) {FALSE} else {TRUE}, weights=NA)

#UNCONNECTED Logical, what to do if the graph is unconnected. If FALSE, the function will
#return a number that is one larger the largest possible diameter, which is always
#the number of vertices. If TRUE, the diameters of the connected components
#will be calculated and the largest one will be returned

#WEIGHTS Optional positive weight vector for calculating weighted distances. If the graph
#has a weight edge attribute, then this is used by default.

```


Mean Distance: The average length of all the shortest paths from or to the nodes in the network.
```{r}
network_stats$mean_distance <- mean_distance(network, directed = TRUE, 
                                   unconnected = if (network_stats$isolates == 0) {FALSE} else {TRUE})

```


Edge Density: ratio of the number of edges and the number of possible edges. The density is in the middle (1 is max, 0 is min density).
```{r}
network_stats$density <- edge_density(network, loops=TRUE) 
```


Transitivity: 
```{r}
network_stats$transitivity <- transitivity(network, weights = TRUE, type = "undirected")

```




### Sources
http://www.people.vcu.edu/~gasmerom/MAT131/graphs.html#:~:text=Two%20vertices%20are%20said%20to,edge%20(arc)%20connecting%20them.&text=Adjacent%20edges%20are%20edges%20that%20share%20a%20common%20vertex.&text=The%20degree%20of%20a%20vertex,edges%20incident%20with%20that%20vertex.&text=A%20graph%20is%20said%20to,are%20joined%20by%20a%20path.

https://github.com/uva-bi-sdad/oss-2020/blob/master/src/github-network-analysis/02_international-collaboration/03_country-networks/03_ctry-nets-cum-analysis.Rmd

https://www.jessesadler.com/post/network-analysis-with-r/

https://kateto.net/wp-content/uploads/2018/03/R%20for%20Networks%20Workshop%20-%20Ognyanova%20-%202018.pdf

https://kateto.net/network-visualization

https://ebookcentral-proquest-com.proxy01.its.virginia.edu/lib/uva/reader.action?docID=4199226&ppg=185

https://lehd.ces.census.gov/data/lodes/LODES7/LODESTechDoc7.4.pdf

https://igraph.org/r/doc/igraph.pdf












