---
title: "Caculating Variables"
output: html_notebook
---

Here I am calculating variables based on geographical data

```{r}
library(tidycensus)
library(tidyverse)
library(viridis)
library(sf)
library(ggthemes)
library(RColorBrewer)
library(ggplot2)
```

Obtain the API key and set install=TRUE during the first time use, to use the same key automatically you run the program 
```{r}
census_api_key("bf083ecaf2c15d0b10906c4948c5759589df1817", install=TRUE)
```

```{r}
var_ACS5 <- load_variables(2018, "acs5", cache = TRUE)
View(var_ACS5)
```

```{r}
vars <- c("B08604_001", 
          "B08101_001",
          "B08101_002",
          "B08101_003",
          "B08101_004",
          "B08101_005",
          "B08101_006",
          "B08101_007",
          "B08101_008",
          "B08603_001",
          "B08603_002",
          "B08603_003",
          "B08603_004",
          "B08603_005",
          "B08603_006",
          "B08603_007",
          "B08603_008",
          "B08603_009",
          "B08603_010",
          "B08603_011",
          "B08603_012",
          "B08603_013",
          "B08541_001",
          "B08541_002",
          "B08601_001",
          "B08601_003",
          "B08601_004",
          "B08601_010",
          "B08301_016",
          "B08301_017",
          "B08301_018",
          "B08301_020",
          "B08301_021",
          "B08301_019",
          "B08528_001",
          "B08528_002",
          "B08528_009",
          "B08528_005",
          "B08528_006",
          "B08528_007",
          "B08528_008",
          "B08528_069",
          "B08526_001",
          "B08526_002",
          "B08526_003",
          "B08526_004",
          "B08526_005",
          "B08526_006",
          "B08526_007",
          "B08526_008",
          "B08526_009",
          "B08526_010",
          "B08526_011",
          "B08526_012",
          "B08526_013",
          "B08526_014",
          "B08524_001",
          "B08524_002",
          "B08524_003",
          "B08524_004",
          "B08524_005",
          "B08524_006",
          "B08524_007"
)
```



When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

