---
title: "Flows into Fairfax"
author: "Owen Hart"
date: "7/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
```


Reading in data tables
```{r}
jobs_all <- read_csv("~/git/dspg20fairfax/data/od/jobsall.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))
jobs_fed <- read_csv("~/git/dspg20fairfax/data/od/jobs_federal.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))
jobs_priv <- read_csv("~/git/dspg20fairfax/data/od/jobs_priv.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))


va_xw <- read_csv("~/git/dspg20fairfax/data/original/va_xwalk.csv")
```


Parsing GEOID into FIPS codes
```{r} 
jobs_all$h_state <- substr(jobs_all$h_geocode_tract, 1, 2)
jobs_all$h_county <- substr(jobs_all$h_geocode_tract, 3, 5)
jobs_all$h_tract <- substr(jobs_all$h_geocode_tract, 6, 11)

jobs_fed$h_state <- substr(jobs_fed$h_geocode_tract, 1, 2)
jobs_fed$h_county <- substr(jobs_fed$h_geocode_tract, 3, 5)
jobs_fed$h_tract <- substr(jobs_fed$h_geocode_tract, 6, 11)

jobs_priv$h_state <- substr(jobs_priv$h_geocode_tract, 1, 2)
jobs_priv$h_county <- substr(jobs_priv$h_geocode_tract, 3, 5)
jobs_priv$h_tract <- substr(jobs_priv$h_geocode_tract, 6, 11)

# Now parsing workplace GEOIDs
jobs_all$w_state <- substr(jobs_all$w_geocode_tract, 1, 2)
jobs_all$w_county <- substr(jobs_all$w_geocode_tract, 3, 5)
jobs_all$w_tract <- substr(jobs_all$w_geocode_tract, 6, 11)

jobs_fed$w_state <- substr(jobs_fed$w_geocode_tract, 1, 2)
jobs_fed$w_county <- substr(jobs_fed$w_geocode_tract, 3, 5)
jobs_fed$w_tract <- substr(jobs_fed$w_geocode_tract, 6, 11)

jobs_priv$w_state <- substr(jobs_priv$w_geocode_tract, 1, 2)
jobs_priv$w_county <- substr(jobs_priv$w_geocode_tract, 3, 5)
jobs_priv$w_tract <- substr(jobs_priv$w_geocode_tract, 6, 11)

```

What % of jobs flow from outside of Fairfax County?
```{r}
# for jobs all (059 is Fairfax code)
bool_cond <-jobs_all$h_county != "059" & jobs_all$w_county == "059"
#this means all jobs where residence is outside of Fairfax but workplace is in Fairfax

flows_in <- jobs_all[bool_cond, ]

#percents <- data.frame("type" = c(1, 2), "shares" = c((mean(bool_cond) * 100), ((1 - mean(bool_cond)) * 100)))
#percents
#ggplot(percents, aes("", y = shares, fill = type)) + geom_bar(stat="identity", width=.2) +
#  coord_polar("y", start=0) 

pie(c(mean(bool_cond) * 100, (1 - mean(bool_cond)) * 100), labels = c("% Jobs in Fairfax flowing from outside", "% Jobs in Fairfax flowing from within"))
```

Of the flows into Fairfax County from outside, what percent are from other counties in Virginia?
```{r}
perc_flows_other_counties_VA <- mean(flows_in$h_state == "51") *100
perc_flows_other_counties_VA

agg_cnty_VA <- flows_in[flows_in$h_state == "51", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))

#Use VA crosswalk to get county names
va_xw$h_county <- substring(va_xw$cty, 3, 6)
essential <- va_xw %>%
  select(h_county, ctyname)

agg_cnty_VA_with_names <- inner_join(x = agg_cnty_VA, y = unique(essential), by = "h_county")

trimmed <- agg_cnty_VA_with_names[order(-agg_cnty_VA_with_names$S000), ][1:10, ]
ggplot(trimmed, aes(ctyname, S000)) + geom_col() + coord_flip() + ggtitle("Number of Jobs flowing into Fairfax from other counties in Virginia")
```


Of the flows into Fairfax County from outside, what percent are from other states?
```{r}
percent_flows_other_states <- mean(flows_in$h_state != "51") * 100
percent_flows_other_states

agg_states_no_names <- flows_in[flows_in$h_state != "51", ] %>%
  group_by(h_state) %>%
  summarise(S000 = sum(S000))

trimmed <- agg_states_no_names[order(-agg_states_no_names$S000), ][1:10, ]
trimmed$State <- c("Maryland", "District of Columbia", "West Virginia", "Pennsylvania", "North Carolina", 
                   "New Jersey", "New York", "Ohio", "South Carolina", "Florida")

ggplot(trimmed, aes(State, S000)) + geom_col() + coord_flip() + ggtitle("Number of Jobs flowing into Fairfax from other states")

#Use other states cross walks to get state names
```

```{r}
#Adding in column to be able to easily match between different state xwalk files
flows_in[flows_in$h_state != "51", ]
flows_in$cty <- substring(flows_in$h_geocode_tract, 1, 5)
flows_in
```


Building same dataframe as flows_in but with County Name, State column added for relevant rows
```{r}
nc_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/nc_xwalk.csv", col_types = cols(cty = col_character()))
dc_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/dc_xwalk.csv", col_types = cols(cty = col_character()))
pa_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/pa_xwalk.csv", col_types = cols(cty = col_character()))
wv_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/wv_xwalk.csv", col_types = cols(cty = col_character()))
md_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/md_xwalk.csv", col_types = cols(cty = col_character()))
oh_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/oh_xwalk.csv", col_types = cols(cty = col_character()))
ny_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/ny_xwalk.csv", col_types = cols(cty = col_character()))
fl_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/fl_xwalk.csv", col_types = cols(cty = col_character()))
nj_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/nj_xwalk.csv", col_types = cols(cty = col_character()))
sc_xw <- read_csv("~/git/dspg20fairfax/data/original/xwalk/sc_xwalk.csv", col_types = cols(cty = col_character()))
```


```{r}
#NC done
essential <- nc_xw %>%
  select(cty, ctyname)
nc_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#dc done
essential <- dc_xw %>%
  select(cty, ctyname)
dc_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#pa done
essential <- pa_xw %>%
  select(cty, ctyname)
pa_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#wv done
essential <- wv_xw %>%
  select(cty, ctyname)
wv_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#md done
essential <- md_xw %>%
  select(cty, ctyname)
md_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#oh done
essential <- oh_xw %>%
  select(cty, ctyname)
oh_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#ny done
essential <- ny_xw %>%
  select(cty, ctyname)
ny_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#fl done
essential <- fl_xw %>%
  select(cty, ctyname)
fl_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#nj done
essential <- nj_xw %>%
  select(cty, ctyname)
nj_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")

#sc done
essential <- sc_xw %>%
  select(cty, ctyname)
sc_names <- inner_join(x = flows_in, y = unique(essential), by = "cty")
```


```{r}
top_ten_with_names <- rbind(nc_names, dc_names, pa_names, wv_names, md_names, oh_names, ny_names, fl_names, nj_names, sc_names)

agg <- top_ten_with_names %>%
  group_by(ctyname) %>%
  summarise(S000 = sum(S000))

trimmed <- agg[order(-agg$S000), ][1:10, ]
ggplot(trimmed, aes(ctyname, S000)) + geom_col() + coord_flip() + ggtitle("Number of Jobs flowing into Fairfax from other\n counties in other states")

```
