---
title: "Flows into Fairfax"
author: "Owen Hart"
date: "7/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(forcats)
library(sf)
library(tmap)
library(leaflet)
library(stplanr)
library(tigris)
```


Reading in data tables
```{r}
jobs_all <- read_csv("~/Git/dspg20fairfax/data/od/jobsall.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))
jobs_fed <- read_csv("~/Git/dspg20fairfax/data/od/jobs_federal.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))
jobs_priv <- read_csv("~/Git/dspg20fairfax/data/od/jobs_priv.csv", col_types = cols(h_geocode_tract = col_character(), w_geocode_tract = col_character()))
va_xw <- read_csv("~/Git/dspg20fairfax/data/original/va_xwalk.csv")
```


Parsing GEOID into FIPS codes
```{r} 
jobs_all$h_state <- substr(jobs_all$h_geocode_tract, 1, 2)
jobs_all$h_county <- substr(jobs_all$h_geocode_tract, 3, 5)
jobs_all$h_tract <- substr(jobs_all$h_geocode_tract, 6, 11)
jobs_all$h_stcty <- substr(jobs_all$h_geocode_tract, 1, 5)

jobs_fed$h_state <- substr(jobs_fed$h_geocode_tract, 1, 2)
jobs_fed$h_county <- substr(jobs_fed$h_geocode_tract, 3, 5)
jobs_fed$h_tract <- substr(jobs_fed$h_geocode_tract, 6, 11)
jobs_fed$h_stcty <- substr(jobs_fed$h_geocode_tract, 1, 5)

jobs_priv$h_state <- substr(jobs_priv$h_geocode_tract, 1, 2)
jobs_priv$h_county <- substr(jobs_priv$h_geocode_tract, 3, 5)
jobs_priv$h_tract <- substr(jobs_priv$h_geocode_tract, 6, 11)
jobs_priv$h_stcty <- substr(jobs_priv$h_geocode_tract, 1, 5)

# Now parsing workplace GEOIDs
jobs_all$w_state <- substr(jobs_all$w_geocode_tract, 1, 2)
jobs_all$w_county <- substr(jobs_all$w_geocode_tract, 3, 5)
jobs_all$w_tract <- substr(jobs_all$w_geocode_tract, 6, 11)
jobs_all$w_stcty <- substr(jobs_all$w_geocode_tract, 1, 5)

jobs_fed$w_state <- substr(jobs_fed$w_geocode_tract, 1, 2)
jobs_fed$w_county <- substr(jobs_fed$w_geocode_tract, 3, 5)
jobs_fed$w_tract <- substr(jobs_fed$w_geocode_tract, 6, 11)
jobs_fed$w_stcty <- substr(jobs_fed$w_geocode_tract, 1, 5)

jobs_priv$w_state <- substr(jobs_priv$w_geocode_tract, 1, 2)
jobs_priv$w_county <- substr(jobs_priv$w_geocode_tract, 3, 5)
jobs_priv$w_tract <- substr(jobs_priv$w_geocode_tract, 6, 11)
jobs_priv$w_stcty <- substr(jobs_priv$w_geocode_tract, 1, 5)
```

What % of jobs flow from outside of Fairfax County?
```{r}
# for jobs all (059 is Fairfax code)
#this means all jobs where residence is outside of Fairfax but workplace is in Fairfax
flows_in <- jobs_all %>% filter(h_stcty != "51059" & w_stcty == "51059")
bool_cond <- jobs_all$h_stcty != "51059" & jobs_all$w_stcty == "51059"

percents <- data.frame("type" = c("% Jobs in Fairfax flowing from outside", "% Jobs in Fairfax flowing from within"), "shares" = c((mean(bool_cond) * 100), ((1 - mean(bool_cond)) * 100)))
percents
pie <- ggplot(percents, aes("", y = shares, fill = type)) +
  geom_bar(stat="identity", width=1) + 
  geom_text(aes(label = paste0(round(shares), "%")), position = position_stack(vjust = 0.5))
 
pie <- pie  +
  coord_polar("y", start=0) +
  scale_fill_manual(values=c("#D1E0BF", "#0E879C")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "Fairfax Jobs Inflows Breakdown") +
  theme_classic() +
  theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(hjust = 0.5))
pie
```

Of the flows into Fairfax County from outside, what percent are from other counties in Virginia?
```{r}
perc_flows_other_counties_VA <- mean(flows_in$h_state == "51") *100
perc_flows_other_counties_VA
agg_cnty_VA <- flows_in[flows_in$h_state == "51", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup()
#Use VA crosswalk to get county names
va_xw$h_county <- substring(va_xw$cty, 3, 6)
essential <- va_xw %>%
  select(h_county, ctyname)
agg_cnty_VA_with_names <- inner_join(x = agg_cnty_VA, y = unique(essential), by = "h_county")
agg_cnty_VA_trimmed <- agg_cnty_VA_with_names[order(-agg_cnty_VA_with_names$S000), ][1:10, ]
agg_cnty_VA_trimmed %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill = "#E57200") +
  ggtitle("Number of Jobs flowing into Fairfax\n from other counties in Virginia") +
  xlab("Number of Jobs") + 
  ylab("County Name") +
  theme_minimal()
```


Of the flows into Fairfax County from outside, what percent are from other states?
```{r}
percent_flows_other_states <- mean(flows_in$h_state != "51") * 100
percent_flows_other_states
agg_states_no_names <- flows_in[flows_in$h_state != "51", ] %>%
  group_by(h_state) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup()
agg_states_trimmed <- agg_states_no_names[order(-agg_states_no_names$S000), ][1:10, ]
agg_states_trimmed$State <- c("Maryland", "District of Columbia", "West Virginia", "Pennsylvania", "North Carolina", "New Jersey", "New York", "Ohio", "South Carolina", "Florida")
agg_states_trimmed %>%
  mutate(ctyname = fct_reorder(State, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "#E57200") +
  ggtitle("Number of Jobs flowing into Fairfax from other states") +
  xlab("Number of Jobs") + 
  ylab("State") +
  theme_minimal()
```


Building same dataframe as flows_in but with County Name, State column added for relevant rows
Takes a long time to run
```{r}
nc_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/nc_xwalk.csv", col_types = cols(cty = col_character()))
dc_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/dc_xwalk.csv", col_types = cols(cty = col_character()))
pa_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/pa_xwalk.csv", col_types = cols(cty = col_character()))
wv_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/wv_xwalk.csv", col_types = cols(cty = col_character()))
md_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/md_xwalk.csv", col_types = cols(cty = col_character()))
oh_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/oh_xwalk.csv", col_types = cols(cty = col_character()))
ny_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/ny_xwalk.csv", col_types = cols(cty = col_character()))
fl_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/fl_xwalk.csv", col_types = cols(cty = col_character()))
nj_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/nj_xwalk.csv", col_types = cols(cty = col_character()))
sc_xw <- read_csv("~/Git/dspg20fairfax/data/original/xwalk/sc_xwalk.csv", col_types = cols(cty = col_character()))
```


```{r}
#NC done
essential <- nc_xw %>%
  select(cty, ctyname)
nc_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#dc done
essential <- dc_xw %>%
  select(cty, ctyname)
dc_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#pa done
essential <- pa_xw %>%
  select(cty, ctyname)
pa_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#wv done
essential <- wv_xw %>%
  select(cty, ctyname)
wv_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#md done
essential <- md_xw %>%
  select(cty, ctyname)
md_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#oh done
essential <- oh_xw %>%
  select(cty, ctyname)
oh_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#ny done
essential <- ny_xw %>%
  select(cty, ctyname)
ny_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#fl done
essential <- fl_xw %>%
  select(cty, ctyname)
fl_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#nj done
essential <- nj_xw %>%
  select(cty, ctyname)
nj_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))

#sc done
essential <- sc_xw %>%
  select(cty, ctyname)
sc_names <- inner_join(x = flows_in, y = unique(essential), by = c("h_stcty" = "cty"))
```


```{r}
top_ten_with_names <- rbind(nc_names, dc_names, pa_names, wv_names, md_names, oh_names, ny_names, fl_names, nj_names, sc_names)
agg_cnty_other_with_names <- top_ten_with_names %>%
  group_by(ctyname) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup()
agg_cnty_other_trimmed <- agg_cnty_other_with_names[order(-agg_cnty_other_with_names$S000), ][1:10, ]
agg_cnty_other_trimmed %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "#E57200") +
  ggtitle("Number of Jobs flowing into Fairfax from other\n counties in other states") +
  xlab("Number of Jobs") + 
  ylab("County Name") +
  theme_minimal()
```

Mapping Flows (have to load in all top 10 states)
```{r}
va_points <- tracts("VA", class = "sf")
md_points <- tracts("MD", class = "sf")
dc_points <- tracts("DC", class = "sf")
wv_points <- tracts("WV", class = "sf")
pa_points <- tracts("PA", class = "sf")
nc_points <- tracts("NC", class = "sf")
nj_points <- tracts("NJ", class = "sf")
ny_points <- tracts("NY", class = "sf")
oh_points <- tracts("OH", class = "sf")
sc_points <- tracts("SC", class = "sf")
fl_points <- tracts("FL", class = "sf")
```

only do states and counties given by top 10 graphs

Doing states first:
(in flows)
```{r}
flows_in_state <- flows_in[flows_in$h_state != "51", ] %>%
  group_by(h_state, w_state) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup() %>%
  arrange(-S000)
#get states names
flows_in_state <- flows_in_state[1:10, ]
flows_in_state$State <- c("Maryland", "District of Columbia", "West Virginia", "Pennsylvania", "North Carolina", "New Jersey", "New York", "Ohio", "South Carolina", "Florida")

states_flow_lines <-st_as_sf(od2line(flows_in_state, states()[3:10]))
tmap_mode("view")
pal <- viridis::viridis(n = 3, option = "C")
top10states_flowsmap <- tm_shape(states_flow_lines %>% select(State, w_state, S000, geometry)) +
  tm_lines(lwd = 3.5, 
           col = "S000", 
           title.col = "Number of Jobs", 
           style = "quantile",
           palette = c("#60999A", "#E57200"),
           popup.vars = c("Number of Jobs" = "S000")) +
  tm_basemap("OpenStreetMap.HOT") +
  tm_view(set.zoom.limits = c(5, 18), basemaps.alpha = 0.5)
top10states_flowsmap
```


Doing VA Counties now:
```{r}
#by VA county
flows_in_VA_cnty <- flows_in[flows_in$h_state == "51", ] %>%
  group_by(h_stcty, w_stcty) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup() %>%
  arrange(-S000)
#join with names of h_county

essential <- va_xw %>%
  select(cty, ctyname)
essential$cty <- as.character(essential$cty)
flows_in_VA_cnty <- inner_join(x = flows_in_VA_cnty, y = unique(essential), by = c("h_stcty" = "cty"))

va_cnty <- counties("VA")
va_cnty_flow_lines <- st_as_sf(od2line(flows_in_VA_cnty[1:10, ], va_cnty, 
                                       origin_code = "h_stcty", dest_code = "w_stcty", 
                                 zone_code = "GEOID"))

tmap_mode("view")
pal <- viridis::viridis(n = 3, option = "C")
top10VA_cnty_flowsmap <- tm_shape(va_cnty_flow_lines %>% select(ctyname, w_stcty, S000, geometry)) +
  tm_lines(lwd = 3.5, col = "S000", title.col = "Number of Jobs", style = "quantile",
           palette = c("#60999A", "#E57200"),
           popup.vars = c("Number of Jobs" = "S000")) +
  tm_basemap("OpenStreetMap.HOT") +
  tm_view(set.zoom.limits = c(8, 18), basemaps.alpha = 0.5)
top10VA_cnty_flowsmap
```

Doing counties outside of VA now:
```{r}
top_ten_with_names$h_stcty <- substr(top_ten_with_names$h_geocode_tract, 1, 5)
top_ten_with_names$w_stcty <- substr(top_ten_with_names$w_geocode_tract, 1, 5)

flows_from_other <- top_ten_with_names %>%
  group_by(h_stcty, w_stcty, ctyname) %>%
  summarise(S000 = sum(S000)) %>%
  ungroup() %>%
  arrange(-S000)
all_points <- rbind(va_points, md_points, dc_points, wv_points, pa_points, nc_points, ny_points, ny_points, oh_points, sc_points, fl_points)
all_points$h_stcty <- paste(all_points$STATEFP, all_points$COUNTYFP, sep = "")

other_cnty_flow_lines <- od2line(flows_from_other[1:10, ], all_points,
                                 origin_code = "h_stcty", dest_code = "w_stcty", 
                                 zone_code = "h_stcty")
other_cnty_flow_lines <- st_as_sf(other_cnty_flow_lines)
other_cnty_flow_lines
tmap_mode("view")
pal <- viridis::viridis(n = 3, option = "C")
top10other_cnty_flowsmap <- tm_shape(other_cnty_flow_lines %>% select(ctyname, w_stcty, S000, geometry)) +
  tm_lines(lwd = 3.5, col = "S000", title.col = "Number of Jobs", style = "quantile",
           palette = c("#60999A", "#E57200"),
           popup.vars = c("Number of Jobs" = "S000")) +
  tm_basemap("OpenStreetMap.HOT") +
  tm_view(set.zoom.limits = c(7, 18), basemaps.alpha = 0.5)
top10other_cnty_flowsmap
```
